---
title: "Class10 –Comparative structure analysis of Adenylate Kinase"
author: "Eric Garcia Ordunez (A16316409)"
format: html
editor: visual
---

# Comparative analysis of ADK

ADK (adenelate Kinase) is an important drug target and we would love to know how it works - i.e. molecular structure 

There has been lots of work done on this protein due to its important including lots of crystals structures.

We will begin with getting an example ADK sequence 

> Q 10: Which of the packages above is found only on BioConductor and not CRAN?

msa

> Q11: Which of the above packages is not found on BioConductor or CRAN?

bio3d-view

> Q12: True or False? Functions from the devtools package can be used to install packages from GitHub and BitBucket?

True

```{r}
library("bio3d")

aa <- get.seq("1ake_A")

aa
```

> Q13. How many amino acids are in this sequence, i.e. how long is this sequence? 

214

We can now run BLAST with this sequence 

```{r}
#b <- blast.pdb(aa)
```

```{r}
#hits <- plot(b)
```

Let's see waht is in our `hits` object

```{r}
#hits$pdb.id
hits <- NULL
hits$pdb.id <- c('1AKE_A','6S36_A','6RZE_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','3GMT_A','4PZL_A')

```

Now we can download all these PDB structure files: 

```{r}
# Download related PDB files
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```


Now I want to align and superpose these structures which are all over the place. 

```{r}
# Align related PDBs
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```

let's have a look at our `pbds` object
```{r}
pdbs
```

Now we have our aligned and superposed we can perform all sorts of analysis on them. Lets do PCA... 

```{r}
# Perform PCA
pc.xray <- pca(pdbs)
plot(pc.xray)
```

Results of PCA on Adenylate kinase X-ray structures. Each dot represents one PDB structure.

We can cluster the structures by RMSD  (or any other method).

```{r}
# Calculate RMSD
rd <- rmsd(pdbs)

# Structure-based clustering
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)
```

This is going to take the PCA and make a movie- also called a trajectory of the major difference (i.e. structural displacements of ADK)

```{r}
# Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```


















